---
title: "Covid91 vaccine study Final2024"
author: "Xavier Cruz"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="finalSumr2024.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

## Males

Is the Covid vaccine effetive in males?

$H_0$: The new preventative vaccine (Covid921) does not exhibit any effectivness on males.  
$H_A$: The new preventative vaccine (Covid921) does exhibit effectivness on male.  

Cat~Cat

```{r}
Males = FinalData[FinalData$sex=='M',]
head(Males)
```
## Graphical Results

```{r}
barchartGC(~infected + treatment,data=Males)
barchartGC(~infected + treatment,data=Males, type="percent")
```
The graph above shows that the vaccine is effective in preventing covid19 in males. The amount of males who received the vaccine and contracted covid19 was lower than the men who were given the placebo.'

## Numerical Results

```{r}
table1 <- xtabs(~infected + treatment, data=Males)
rowPerc(table1)
colPerc(table1)
```

The table above shows that only 26.20% of men that were given the vaccine and contracted covid-19, while 73.80% of men those who received the placebo contracted covid-19.

## Inferential Results  
```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

Because the p-value from the chi-squared test is 0.00000000000000022, that means that is less than 0.05, we must reject the null hypothesis. The odds ratio from fisher's test was 2.815 which means that the vaccine has a 2.8 chance of being effective in males.  


## Females

## Methods 
$H_0$: The preventative vaccine does not exhibit its effective in females        
$H_A$: The preventative vaccine exhibits its effective in females.  
Cat~Cat. 

```{r}
Females = FinalData[FinalData$sex=='F',]
head(Females)
```

## Graphical Results

```{r}
barchartGC(~infected + treatment,data=Females)
barchartGC(~infected + treatment,data=Females, type="percent")
```

The bar graph shows that the vaccine is more effective in women who contracted covid-19 compared to women who were given the placebo.

## Numerical Results  
```{r}
table1 <- xtabs(~infected + treatment, data=Females)
rowPerc(table1)
colPerc(table1)
```

The numerical results shows that about 35.07% of women who received the vaccine contracted covid-19. However, the women who contracted covid-19 received the placebo which was 64.93% which implies that the vaccine is effective in preventing covid-19 in women.   

## inferential Results   
```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The p-value from the chi-squared test is 0.00000001914 which is less than 0.05 therefore we reject the null hypothesis. The odds ratio from the fisher's exact test is 1.856 which means that there is approximately a 1.86 chance that the vaccine will be effective in women.



## LGBTQ

## Methods 

$H_0$ The preventative vaccine does not exhibit its effective in people who are LGBTQ.        
$H_A$ The preventative vaccine exhibits its effective in people who are LGBTQ.    
Cat~Cat

```{r}
gay = FinalData[FinalData$LGBTQ=='gay',]
head(gay)
```


## Graphical Results 

```{r}
barchartGC(~infected + treatment,data=gay)
barchartGC(~infected + treatment,data=gay, type="percent")
```

The bar graph shows that the vaccine is effective in preventing covid-19 in people who are LGBTQ. There is about 20% of LGBTQ people who contracted covid-19 and received the vaccine which is significantly lower than the amount of LGBTQ people who received the placebo and contracted covid-19. 


## Numerical Results  
```{r}
table1 <- xtabs(~infected + treatment, data=gay)
rowPerc(table1)
colPerc(table1)
```

The table shows that 20% of LGBTQ people who received the vaccine also contracted covid-19. Compared to the LGBTQ people who received the placebo and contracted vaccine which is 80%, we can assume the vaccine is effective at preventing LGBTQ people from contracting covid-19.       

## Infrential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The p-value from the chi-squared test is 0.000002266 which is less than 0.05 therefore we reject the null hypothesis. The odds ratio from the fisher's exact test is 4.27 which means that the vaccine has a 4.27 chance of being effective in preventing covid19 in people who are LGBTQ.   


## Druggies

## Methods  

$H_0$ The preventative vaccine does not exhibit its effective in people who are drug users.    
$H_A$ The preventative vaccine exhibits its effective in people who are drug users.  
Cat~Cat


```{r}
DrugUser = FinalData[FinalData$DrugUser=='yes',]
```

## Graphical Results    
```{r}
barchartGC(~infected + treatment,data=DrugUser)
barchartGC(~infected + treatment,data=DrugUser, type="percent")
```
The bar graphs shows that vaccine is more likely to be less effective in people who are drug users.The amount of people who received the vaccine and contracted covid19 passed 80% which is significantly higher than drug users who received the placebo. 

## Numerical Results  

```{r}
table1 <- xtabs(~infected + treatment, data=DrugUser)
rowPerc(table1)
colPerc(table1)
```

The table shows that 91.67% of drug users who contracted covid-19 also received the vaccine. This implies that the preventative vaccine is not effective in people who are drug users because the amount of drug users who contracted covid-19 and received the placebo was 8.33%.

## Inferential Results      

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The p-value from the chi-squared test is 0.0000000001414 which is less than 0.05 therefore we reject the null hypothesis. The odds ratio from the fisher's exact test is 0.087 which means that the vaccine has a lower chance of being effective in people who are drug users.



# Overall Results and Conclusions

## Males     
From the data gathered we can conclude that the vaccine is effective in preventing covid-19 in males. The numerical table shows that 26.20% men contracted covid-19 and received the vaccine compared to men who were given the placebo which was 73.80%. The p-value from the chi-squared was 0.00000000000000022 which is less than 0.05 therefore we reject the null hypothesis. The odds of the vaccine being effective in men is about a 2.8 chance, so we can assume that the vaccine will be effective in preventing covid19 in males.  

##Females      
From the data we gathered, we can conclude that the vaccine is effective in preventing covid-19 in females. The numerical tables shows that 35.07% women who contracted covid-19 received the vaccine compared to the women who received the placebo which was 64.93%. The p-value from the chi-squared test was 0.00000001914 which is less than 0.05 therefore we reject the null hypothesis. The odds of the vaccine being effective in women is about a 1.86 chance. We can assume that the vaccine is effective in preventing covid-19 in females based on the data.      

##LGBTQ       
With the data we gathered, we can conclude that the vaccine exhibits effectiveness in preventing covid19 in people who are LGBTQ. The numerical tables shows that 20% of people who are LGBTQ contracted covid-19 and also received the vaccine. The number of LGBTQ people who received the placebo was 80% which means the vaccine is effective in preventing covid19 in people who are LGBTQ. The p-value from the chi-squared test is 0.000002266 which is less than 0.05 therefore we reject the null hypothesis. The odds ratio from the fisher's exact test is 4.27 which means the vaccine has a 4.27 chance of being effective in preventing covid19 in people who are LGBTQ. It is safe to assume that the new preventative vaccine will be effective in people who are LGBTQ. 

##Drug Users      
With the data we gathered, we can conclude that the vaccine is less likely to be effective in preventing covid-19 in drug users. In the numerical table, approximately 91.67% of drug users who contracted covid-19 also received the vaccine compared to drug users who received the placebo which was 8.33%. The p-value from the chi-squared test was 0.0000000001414 which is less than 0.05 therefore we reject the null hypothesis. The odds ratio from the fisher's exact test is 0.087 which means that the vaccine is less likely to effective in people who are drug users. From these results we can assume that the vaccine will not be effective in preventing covid19 in people who are drug users.   
